---
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE,
          fig.width = 5,
          fig.height = 3,
          out.width = "75%",
          fig.align = "center")
options(list(dplyr.summarise.inform = FALSE))
theme_set( theme_classic() )
```

# Package to fit 2D RDD models

This package fits 2D RDD models using loess or gaussian processes. The package is still in development and is not yet on CRAN.


## Installation

You can install the development version of twodrdd from [GitHub](https://github.com/lily-an1/twodrdd) with:
```
# install.packages("devtools")
devtools::install_github("lily-an1/twodrdd")
```


## Example

This is a basic example which shows you how to fit a 2D RDD model:

```{r}
library( twodrdd )
dat = gen_dat_sim( sim = 5, n = 700, rho = 0.80, s = 1 )
res <- analysis(dat$data, 
                n_sentinel = 20,
         include_BINDING = TRUE,
         include_FRONTIER = TRUE,
         include_OLS = TRUE,
         include_GP = TRUE,
         include_GP_RES = TRUE,
         include_loess = TRUE,
         include_PARA_LINEAR = TRUE,
         include_PARA_QUAD = TRUE
         )

knitr::kable( res, digits = 2)
```

There are other functions to get more rich descriptions of the analysis.  See, e.g., `gaussianp()`.



